Three 38 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-03-14 papers/authors.tex
compile paper
update ms

\author[0000-0002-0514-5538]{L. G. Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
%\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

% Key authors:
% ... Kinematics
\author[0000-0002-6549-9792]{R.~Kerr}
\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations

% ... stellar rotation & the initial crossmatch
\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}

% \author[0000-0003-1298-9699]{K. Masuda}
% \affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}
% 
% % ... planet-fitting, kinematics, meta.
\author{L. A. Hillenbrand}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
% 
% % ... RM fitting
% \author[0000-0001-7409-5688]{G. Stefansson}
% \altaffiliation{Henry Norris Russell Fellow}
% \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%
% PFS Collaborators
%
\author[0000-0001-8638-0320]{A. W. Howard}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
%
\author[0000-0002-0531-1073]{H. Isaacson}
\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
% pending

% AO IMAGING
\author[0000-0001-9811-568X]{A.~L.~Kraus}
\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations

>>>>

\author[0000-0002-0514-5538]{L. G. Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}

% Key authors:
% ... Kinematics
\author[0000-0002-6549-9792]{R.~Kerr}
\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
%
% ... stellar rotation & the initial crossmatch
\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
%
% HIRES
\author[0000-0002-0531-1073]{H. Isaacson}
\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
%
% planet-fitting, kinematics, cluster.
\author{L. A. Hillenbrand}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
%
% HIRES Collaborators
\author[0000-0001-8638-0320]{A. W. Howard}
\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
%
% AO IMAGING
\author[0000-0001-9811-568X]{A.~L.~Kraus}
\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
%
% TRES
\author[0000-0001-6637-5401]{A. Bieryla}
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
%
% TRES
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
%
% AO / NIRC2
\author[0000-0003-0967-2893]{E. A.~Petigura}
\affiliation{Department of Physics \& Astronomy, University of California Los Angeles, Los Angeles, CA 90095, USA}
%
% AO / NIRC2
\author[0000-0001-8832-4488]{D. Huber}
\affiliation{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}

### 2022-03-14 papers/ms.tex
update ms

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of stellar cluster evolution.
  Here we present an analysis of a group of $\approx$40 million
  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.  The group contains
  four known Kepler Objects of Interest:
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
  Kepler-1627 is a Neptune-sized planet in a sub-component of the
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.  Here we focus on the latter three
  systems, which are in other sub-components of the complex (RSG-5 and
  CH-2).  Based on kinematic evidence from Gaia, stellar rotation
  periods from TESS, and spectroscopy, these three systems are also
  $38 \pm 6$ million years old.  Based on the transit shapes and high
  resolution imaging, they are all most
  likely planets (false positive probabilities of $6\times10^{-9}$,
  $5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
  and KOI-7913 respectively).  Kepler-1643 and KOI-7913 are therefore
  the first empirical demonstration that mini-Neptunes with sizes of
  $\approx$2 Earth radii exist at ages of roughly 40 million years.
\end{abstract}

@@

The discovery of young transiting planets is a major frontier in
current exoplanet research.  The reason is that the properties of
these planets can help constrain the timescales for processes
including hot Jupiter migration (CITE), the cooling and contraction of
mini-Neptunes (CITE Rizzuto), the early evolution of close-in
multiplanet systems (CITE David), and whether and how mass-loss
explains the radius valley (e.g., CITE, CITE, Lee+21 on theory; CITE
CITE CITE on observations).

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Galactic positions  and tangential velocities of stars in the
    complex.}  {\bf This is a placeholder before we get real kinematics.}
    Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
    worryingly diffuse ``CH-2''.
		\label{fig:XYZvtang}
	}
\end{figure*}

Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
the Sun, above the galactic plane, and spanning galactic longitudes of
roughly $40^\circ$ to $100^\circ$ degrees.  A direct query of the Gaia
archive imposing these constraints yields $\approx3.4\times10^7$
sources.  Imposing the cuts suggested by \citet[][Appendix
B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
the number of visibility periods used, the astrometric $\chi^2$ of the
single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
$\approx5.0\times10^6$ sources.

@@

In XYZ, and in tangential velocity space, these five thousand sources
are visually clustered in between a few to a dozen groups, depending
on the characteristic size and velocity scales of interest.  The first
suggestion that these objects could all be part of a monolithic
structure, to our knowledge, came from the manual grouping that was
performed between them by \citet{KounkelCovey2019} when defining the
structure they dubbed ``Theia~73''.  Different subsets of these stars
have however previously been noticed and studied.  They include the
$\delta$ Lyr cluster, also known as Stephenson-1
\citep{stephenson_possible_1959}, which comprises a major grouping
around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
v_b)=(-4.5{\rm \kms},-4{\rm \kms})$.  Another individual component
that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
$(v_{l'}, v_b)=(+5.5{\rm \kms},-3.5{\rm \kms})$.  Most of the other
subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
small or dispersed to have previously been analyzed in great detail.

>>>>

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of how stellar clusters evolve.
  Here we present an analysis of a group of $\approx$40 million
  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.  The group contains
  four known Kepler Objects of Interest:
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
  Kepler-1627 is a Neptune-sized planet in a sub-component of the
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.  Here we focus on the latter three
  systems, which are in other sub-components of the complex (RSG-5 and
  CH-2).  Based on kinematic evidence from Gaia, stellar rotation
  periods from TESS, and spectroscopy, these three systems are also
  $38 \pm 6$ million years old.  Based on the transit shapes and high
  resolution imaging, they are all most
  likely planets (false positive probabilities of $6\times10^{-9}$,
  $5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
  and KOI-7913 respectively).  Kepler-1643 and KOI-7913 are therefore
  the first empirical demonstration that mini-Neptunes with sizes of
  $\approx$2 Earth radii exist at ages of roughly 40 million years.
\end{abstract}

@@

The discovery of young transiting planets is a major frontier in
current exoplanet research.  The reason is that the properties of
these planets can help constrain the timescales for processes
including hot Jupiter migration (CITE), the cooling and contraction of
planets with massive gaseous envelopes (CITE Rizzuto), the early
evolution of close-in multiplanet systems (CITE David), and whether
and how mass-loss explains the radius valley (e.g., CITE, CITE, Lee+21
on theory; CITE CITE CITE on observations).

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Galactic positions  and tangential velocities of candidate members of the
    Cep-Her complex.}
	{\it Top row}: On-sky positions in galactic coordinates.
	Black points are stars for which membership is more likely than for
	gray points (``weight'' cuts above 0.10 and 0.02 respectively; see text).
	The $\delta$ Lyr cluster, at $\{ l, b\} \approx \{ 66^\circ, 12^\circ\}$, contains Kepler-1627  \citep{bouma_kep1627_2022}.
	{\it Middle row}:
	Galactic positions.
	The Sun is at $\{X, Y, Z\} = \{-8122, 0, 20.8\}$ pc.
	{\it Bottom row}:
	Galactic tangential velocities.
	The spatial and kinematic overlap between RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc)
	and Kepler-1643 is fairly obvious.
	KOI-7368 and KOI-7913 however are candidate members of a diffuse population,
	that we dub ``CH-2''.
	Although CH-2 appears close to RSG-5 when projected in $(X,Y)$ and $(v_b, v_{l*})$,
	it is spatially  distinct in $(Z,Y)$.
	\label{fig:XYZvtang}
	}
\end{figure*}

\subsection{General Context} % might cut!

Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
the Sun, above the galactic plane, and spanning galactic longitudes of
roughly $40^\circ$ to $100^\circ$ degrees.  Querying the Gaia archive
and imposing the cuts suggested by \citet[][Appendix
B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
the number of visibility periods used, the astrometric $\chi^2$ of the
single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
$\approx5.0\times10^6$ sources.

@@

In XYZ, and in tangential velocity space, these five thousand sources
are visually clustered in between a few to a dozen groups, depending
on the characteristic size and velocity scales of interest.  The first
suggestion that these objects could all be part of a monolithic
structure, to our knowledge, came from the manual grouping that was
performed between them by \citet{KounkelCovey2019} when defining the
structure they dubbed ``Theia~73''.  Different subsets of these stars
have however previously been noticed and studied.  They include the
$\delta$ Lyr cluster, also known as Stephenson-1
\citep{stephenson_possible_1959}, which comprises a major grouping
around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
v_b)=(-4.5{\rm kms},-4{\rm kms})$.  Another individual component
that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
$(v_{l'}, v_b)=(+5.5{\rm kms},-3.5{\rm kms})$.  Most of the other
subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
small or dispersed to have previously been analyzed in great detail.

\subsection{Actual Membership Selection} % might cut!

% TODO: add words to acknowledge that Cep-Her is not yet demonstrated
% to be a thing

To identify stars in this region of space that could be members of the
Cep-Her complex, we follow the methods of \citet{Kerr2021}.  The
analysis therefore broadly consists of two steps.  The first step is
to select stars that photometrically appear young based on the Gaia
EDR3 magnitudes $\{G, G_{\rm RP}, G_{\rm BP}\}$, in combination with
the parallaxes and reddening estimates.  This method works best for
low-mass stars due to their long pre-main-sequence contraction
timescale, and for massive stars because of their fast evolutionary
timescale.  The second step in the analysis is to use the
photometrically-identified young stars as the ``seed'' for a candidate
population in a spatial and kinematic clustering analysis in $\{ X, Y,
Z, c v_b, c v_{l^*} \} $.  The latter step helps identify F, G, and
K-type stars that cannot be identified as young based solely on their
positions in color versus absolute magnitude.

%FIXME TODO TODO TODO: this analysis!
